---
title: "main"
author: "123"
date: "2021/1/11"
output: html_document
---

# Revealing mismatched spatial aggregation of physical and social vitality in London

load packages 
```{r message=FALSE}
library(here)
library(tidyverse)
library(fs)
library(plyr)
library(sf)
library(tmap)
library(tmaptools)
library(janitor) 
library(spatstat) 
library(spdep)
library(ggplot2)
library(rgdal)
library(stringr)
library(patchwork)
here::here()
```

## Data Prepare

* Download the LSOA boundary

```{r eval=FALSE, include=TRUE}
#download.file("https://data.london.gov.uk/download/statistical-gis-boundary-files-london/9ba8c833-6370-4b11-abdc-314aa020d5e0/statistical-gis-boundaries-london.zip",destfile="data/statistical-gis-boundaries-london.zip")
```

* Download London Building Heights data from [emu-analytics](https://www.emu-analytics.com/products/datapacks.php), and online visualization can be found at <https://buildingheights.emu-analytics.net/>
* Download London Point of interest data from [Digimap](https://digimap.edina.ac.uk/)
* Unzip all files
```{r eval=FALSE, include=TRUE}
#listfiles<- list.files(path = here::here("data"), pattern = "*.zip",
#    recursive = TRUE, full.names = TRUE) %>% 
#    lapply(., unzip, exdir = here::here("data")) 
```
* read the data and check
    * the building height data was divided into residence and none residence, we use rbind to merge the data into all building heights, the volume column is the building area for each building
```{r warning=FALSE}
## building Height 
p_buildH_Res <- here::here("data","London_Res.shp")
sf_buildH_Res <- st_read(p_buildH_Res) %>% 
    clean_names() %>% 
    st_transform(., 27700) 

p_buildH_unRes <-  here::here("data","London_NonRes.shp")
sf_buildH_unRes <- st_read(p_buildH_unRes) %>% 
    clean_names() %>% 
    st_transform(., 27700) ## reprojection 

sf_buildH <- rbind(sf_buildH_Res, sf_buildH_unRes)

sf_buildH
```

```{r echo=TRUE}
## lsoa
p_bodry <- here::here("data","LSOA_2011_London_gen_MHW.shp")

sf_LSOA <- st_read(p_bodry) %>% 
    clean_names() %>% 
    st_transform(., 27700) 
```

* filter poi category of small catering business in all pois
```{r}
## poi
p_poi <- here::here("data","poi2020_09.csv")

l_small_catering<-c(1020013,#Cafes, snack bars and tea rooms
                    1020018,#Fast food and takeaway outlets
                    1020019,#Fast food delivery services
                    1020020,#Fish and chip shops
                    1020034,#Pubs, bars and inns
                    1020043 #Restaurants
                    )

df_poi <- read_csv(p_poi) %>%
    clean_names() %>% 
    filter(pointx_class %in% l_small_catering)

sf_poi <- st_as_sf(df_poi,coords = c("feature_easting", "feature_northing"), 
                   crs = 27700)

tm_shape(sf_poi) + 
    tm_dots()
```

## Calculate metrics

### Built-environment intensity 

to calculate the building area in each lsoa, I use `st_join` to join the two table and group by lsoa to sum the volume.
```{r echo=TRUE}
join <- st_join(sf_LSOA, sf_buildH, join = st_intersects)

vol <-  aggregate(list(ResVol = join$volume), 
                  by=list(lsoa11cd =join$lsoa11cd ), 
                  FUN=sum)
```


```{r echo=TRUE}
sf_res <- left_join(sf_LSOA,vol,by=c("lsoa11cd" = "lsoa11cd")) %>% 
    mutate(area=st_area(.) %>% as.vector()) %>% 
    mutate(BV = ResVol/area)

tm_shape(sf_res) + 
  tm_fill("BV", style = "fisher", palette = "Reds") +
  tm_borders(alpha = 0.1) +
  tm_layout(main.title = "Built-environment vitality", main.title.size = 0.7 ,
            legend.position = c("right", "bottom"), legend.title.size = 0.8)+
    tm_scale_bar(position = c("left", "bottom")) + 
    tm_compass(type = "4star", size = 2, position = c("right", "top"))
```

### Social vitality
```{r echo=TRUE}
join_poi <- st_join(sf_LSOA,sf_poi,join = st_intersects)

poi_count <- count(join_poi$lsoa11cd)

sf_result <- left_join(sf_res,poi_count,by=c("lsoa11cd" = "x")) %>% 
    mutate(SV = freq/area)

tm_shape(sf_result) + 
  tm_fill("SV", style = "fisher", palette = "PuBu") +
  tm_borders(alpha = 0.1) +
  tm_layout(main.title = "Social vitality", main.title.size = 0.7 ,
            legend.position = c("right", "bottom"), legend.title.size = 0.8)+
    tm_scale_bar(position = c("left", "bottom")) + 
    tm_compass(type = "4star", size = 2, position = c("right", "top"))
```

### Distribution and correlation 
```{r}
sf_result
```

```{r echo=TRUE, warning=FALSE}
p1 <- ggplot(sf_result, aes(x=BV, fill=sex)) +
  geom_histogram(fill="white", color="black",bins = 40)+
  geom_vline(aes(xintercept=mean(BV)), color="blue",linetype="dashed")+
  xlim(c(0, 15))+
  labs(x="Physical built-environment vitality", y = "Count")+
  theme(text=element_text(family="serif"))

p2 <- ggplot(sf_result, aes(x=SV, fill=sex)) +
  geom_histogram(fill="white", color="black",bins = 40)+
  geom_vline(aes(xintercept=mean(SV)), color="blue",linetype="dashed")+
  xlim(c(0, 0.0003))+
  labs(x="Social vitality", y = "Count")+
  theme(text=element_text(family="serif"))

p1+p2
```
```{r echo=TRUE, message=FALSE}
png(
    filename = "Hist.png",
    type = "cairo",
    res = 300,
    width = 2500, height = 1200,
    bg = "transparent"
)
p1+p2
dev.off()
```



## Spatial Autocorrelation

```{r echo=TRUE, warning=FALSE}
# bivariate LISA
## reference: 
## https://stackoverflow.com/questions/45177590/map-of-bivariate-spatial-correlation-in-r-bivariate-lisa
## https://gist.github.com/rafapereirabr/5348193abf779625f5e8c5090776a228

#======================================================
# Programming some functions

# Bivariate Moran's I
moran_I <- function(x, y = NULL, W){
        if(is.null(y)) y = x

        xp <- (x - mean(x, na.rm=T))/sd(x, na.rm=T)
        yp <- (y - mean(y, na.rm=T))/sd(y, na.rm=T)
        W[which(is.na(W))] <- 0
        n <- nrow(W)

        global <- (xp%*%W%*%yp)/(n - 1)
        local  <- (xp*W%*%yp)

        list(global = global, local  = as.numeric(local))
}


# Permutations for the Bivariate Moran's I
simula_moran <- function(x, y = NULL, W, nsims = 1000){

        if(is.null(y)) y = x

        n   = nrow(W)
        IDs = 1:n

        xp <- (x - mean(x, na.rm=T))/sd(x, na.rm=T)
        W[which(is.na(W))] <- 0

        global_sims = NULL
        local_sims  = matrix(NA, nrow = n, ncol=nsims)

        ID_sample = sample(IDs, size = n*nsims, replace = T)

        y_s = y[ID_sample]
        y_s = matrix(y_s, nrow = n, ncol = nsims)
        y_s <- (y_s - apply(y_s, 1, mean))/apply(y_s, 1, sd)

        global_sims  <- as.numeric( (xp%*%W%*%y_s)/(n - 1) )
        local_sims  <- (xp*W%*%y_s)

        list(global_sims = global_sims,
             local_sims  = local_sims)
}
```


```{r echo=TRUE, warning=FALSE}
#======================================================
# Variables to use in the correlation: white and black population in each census track
x <- sf_result$BV
y <- sf_result$SV

#======================================================
# Adjacency Matrix (Queen)
## reference : https://gis.stackexchange.com/questions/355300/poly2nb-spdep-not-identifying-neighbours-correctly
overlapmat <- st_overlaps(sf_result,sparse=FALSE)
lw <- mat2listw(overlapmat)
W  <- as(lw, "symmetricMatrix")
W  <- as.matrix(W) #[1:nrow(W),1:ncol(W),drop=FALSE]
W  <- W/rowSums(W) ## rowSum auto compile 
W[which(is.na(W))] <- 0

plot(sf_result$geometry)
plot(lw, st_coordinates(st_centroid(sf_result)),col="red",add=TRUE)
```


```{r echo=TRUE, warning=FALSE}
#======================================================
# Calculating the index and its simulated distribution for global and local values

m  <- moran_I(x, y, W)
m[[1]] # global value
```

```{r echo=TRUE}
m_i <- m[[2]]  # local values

local_sims <- simula_moran(x, y, W)$local_sims

# Identifying the significant values 
alpha <- .05  # for a 95% confidence interval
probs <- c(alpha/2, 1-alpha/2)
intervals <- t( apply(local_sims, 1, function(x) quantile(x, probs=probs)))
sig        <- ( m_i < intervals[,1] )  | ( m_i > intervals[,2] )

#======================================================
# Preparing for plotting
sf_result     <- st_as_sf(sf_result)
sf_result$sig <- sig

# Identifying the LISA patterns
xp <- (x-mean(x))/sd(x)
yp <- (y-mean(y))/sd(y)

patterns <- as.character( interaction(xp > 0, W%*%yp > 0) ) 
patterns <- patterns %>% 
        str_replace_all("TRUE","High") %>% 
        str_replace_all("FALSE","Low")
patterns[sf_result$sig==0] <- "Not significant"
sf_result$patterns <- patterns

# Rename LISA clusters
sf_result$patterns2 <- factor(sf_result$patterns, levels=c("High.High", "High.Low", "Low.High", "Low.Low", "Not significant"), labels=c("High BV - High SV", "High BV - Low SV", "Low BV - High SV","Low BV - Low SV", "Not significant"))

# Plotting
ggplot() + 
    geom_sf(data=sf_result, aes(fill=patterns2), color="NA") +
    scale_fill_manual(values = c("orange", "red", "green", "light blue", "grey95")) + 
    guides(fill = guide_legend(title="LISA clusters")) +
    theme_minimal()
```

```{r echo=TRUE}
## st_write(sf_result, here::here("plots","res.shp"))
```

